---
title: "Reproducible Research Project 1"
author: "AbdulRafaeMohammed"
date: "May 16, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# Project Week-2 Reproducible Research

#Task-1
##Loading the Data
```{r, echo=TRUE}
if(!file.exists("activity.csv")){
  t <- tempfile()
  download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",t)
  unzip(t)
  unlink(t)
}

activity <- read.csv("activity.csv")
```


#Task-2
##Calculating Number of Steps in a Day

```{r, echo=TRUE}
steps_per_day <- aggregate(steps ~ date,activity,FUN=sum,na.rm=TRUE)
```


## Plotting Histogram for Number of Steps on Each Day

```{r,echo=TRUE}
hist(steps_per_day$steps, main = "Total Steps on Each Day", col = "orange",xlab = "Number of Days",ylab = "Number of Steps")
```

## Calculating the Mean and Median of Total Number of Steps Taken per Day
```{r,echo=TRUE}
Mean_Steps <- mean(steps_per_day$steps)
Median_steps <- median(steps_per_day$steps)
```

#### The Mean of Number of Steps is : ```r Mean_Steps```
#### The Median of Number of Steps is : ```r Median_steps```

#Task-3
## Calculate the average activity pattern each day.

```{r, echo=TRUE}
avg_steps_per_day <- aggregate(steps ~ interval,activity,FUN=mean,na.rm=TRUE)
```

## Plot the Average Daily Activity Pattern

```{r,echo=TRUE}
plot(avg_steps_per_day$interval, avg_steps_per_day$steps, type ='l',main = "Average Steps on Each Day by Interval", col = "orange",xlab = "Interval",ylab = "Number of Steps")
```

## Calculate the Interval with the maximum steps
```{r,echo=TRUE}
interval_for_max_steps <- avg_steps_per_day[which.max(avg_steps_per_day$steps),1]
```

#### The interval with maximum number of steps is : ```r interval_for_max_steps```

##Task-4
## Impute Missing Values

### Count Number of NA Values

```{r, echo=TRUE}
na_count <- length(which(is.na(activity$steps)))
```

#### The Number of NA values are : ```r na_count```

### Fill in NA Values

```{r, echo=TRUE}
Imputed_Activity_Data <- merge(activity,avg_steps_per_day,by="interval",suffixes = c("",".interval.mean"))

nas <- is.na(Imputed_Activity_Data$steps)

Imputed_Activity_Data$steps[nas] <- Imputed_Activity_Data$steps.interval.mean[nas]


total_steps_per_day_imputed <- aggregate(steps ~ date,Imputed_Activity_Data, sum)
```

### Plot the histogram with imputed steps per day

```{r ,echo=TRUE}
hist(total_steps_per_day_imputed$steps, main="Total Number of Steps Per Day",col= "orange",xlab="Number of Steps",ylab = "Count")
```


### Calculate the mean and median of the imputed values.

```{r,echo=TRUE}
Mean_imputed_Steps <- mean(total_steps_per_day_imputed$steps)
Median_imputed_steps <- median(total_steps_per_day_imputed$steps)
```

#### The Mean of Number of Steps is : ```r Mean_imputed_Steps```
#### The Median of Number of Steps is : ```r Median_imputed_steps```

### Calculate the Diff between the Imputed and Non-Imputed Data

```{r,echo=TRUE}
diff_mean <- Mean_imputed_Steps - Mean_Steps
diff_median <- Median_imputed_steps - Median_steps
```


#### The Difference between the Means of Number of Steps is : ```r diff_mean```
#### The Difference between the Median of Number of Steps is : ```r diff_median```


### Calculate the Diff between the Imputed and Non-Imputed Data

```{r,echo=TRUE}
total_diff_steps <- sum(total_steps_per_day_imputed$steps) - sum(steps_per_day$steps)
```

#### The Difference between the Number of Steps of imputed and non-imputed data is : ```r total_diff_steps```

## Task-5


```{r,echo=TRUE}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
Imputed_Activity_Data$dow = as.factor(ifelse(is.element(weekdays(as.Date(Imputed_Activity_Data$date)),weekdays), "Weekday", "Weekend"))

avg_imputed_steps_per_day <- aggregate(steps ~ interval + dow, Imputed_Activity_Data, mean)

library(lattice)

xyplot(avg_imputed_steps_per_day$steps ~ avg_imputed_steps_per_day$interval|avg_imputed_steps_per_day$dow, main="Average Number of Steps per Day by Interval",xlab="Interval", ylab="Number of Steps",layout=c(1,2), type="l",col="orange")
```

